Rates of true polar wander in convecting planets

### 2014-02-25 tpw_rate.tex
Start adding text about rotational dynamics

@@

@@

which shows the connection between the degree-two component of the density field and the principle moments of inertia.  This may be seen as a form of MacCullagh's formula \citep{stacey1977physics}

>>>>

We may write the nonlinear Liouville equations 
\begin{equation}
{\bf \Omega} \times {\bf \dot{\Omega} } = \frac{1}{\varepsilon I_0 \tau} {\bf \Omega} \times \left( {\bf E \cdot \Omega} \right)
\end{equation}

introducing a unit vector $\mitbf{\omega} = {\bf \Omega}/\|{\bf \Omega} \|$ we may solve this equation for ${\bf \dot{\mitbf{\omega}} }$:
\begin{equation}
 \dot{\mitbf{\omega}}  = \frac{1}{\varepsilon I_0 \tau} \left[ {\bf E \cdot \mitbf{\omega}} - \left( {\bf \mitbf{\omega} \cdot E \cdot \mitbf{\omega} } \right) \mitbf{\omega} \right]
\end{equation}

Note that the quantity in brackets is identical in form to the shear stress on a plane in classical elastostatics.
If we enter the coordinate system of the convective moment of inertia $\bf E$ with principal moments $\lambda_1 \le \lambda_2 \le \lambda_3$ and define the orientation of $\mitbf{\omega}$ with colatitude $\theta$ and longitude $\phi$ we can write this in a more illuminating form (after some tedious algebra):
\begin{equation}
\begin{aligned}
\dot{\theta} &= - \frac{1}{2 \varepsilon I_0 \tau} \sin{2 \theta} \left[ (\lambda_3-\lambda_1) \cos^2{\phi} + (\lambda_3-\lambda_2) \sin^2{\phi} \right] \\
\dot{\phi} &= \;\; \frac{1}{2 \varepsilon I_0 \tau} \sin{2 \phi} \; (\lambda_2 - \lambda_1)
\end{aligned}
\label{milankovitch}
\end{equation}

These equations are a version of what has been called the ``Milankovitch theorem'' \citep{munk1960rotation}.  
Written this way, it is clear that the important quantities are $\theta$, $\phi$, and the differences of the eigenvalues of the convective moment $\bf E$.
Thus it remains to derive estimates for these.

@@

The nonhydrostatic moment of inertia is defined
\begin{equation}
J_{ij} = \int_{V_S} \rho \left( r_q r_q \delta_{ij} - r_i r_j \right) 
\end{equation}

where $V_S$ is the reference spherical volume of the mantle.
If we substitute in Equation \ref{eos} we find

\begin{equation}
J_{ij} = I_0 \delta_{ij} + \int_{V_S} \rho_0 \alpha T \left( r_q r_q \delta_{ij} - r_i r_j \right) = I_0 \delta_{ij} + E_{ij} 
\end{equation}

Nondimensionalizing the integral for $E_ij$ we find

\begin{equation}
J_{ij} = I_0 \delta_{ij} + \int_{V_S^\prime} \alpha T^\prime \left( r_q^\prime r_q^\prime \delta_{ij} - r_i^\prime r_j^\prime \right) 
\end{equation}

@@

which shows the connection between the degree-two component of the density field and the principle moments of inertia.  This may be seen as a form of MacCullagh's formula \citep{stacey1977physics}.
